### 2016-07-29 experiments/e2/archive/plot_transformations_against_normal_equivalents.Rmd
archiving old e2 Rmd files

---
title: "Experiment 4 of 4"
author: "Matt Green"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---  
# Experiment 4

Each subject saw 192 trials, being 3 cycles of 64 trials. There were 40 subjects. There were 7680 rows in the full data.

# Trial Procedure
First an instruction was presented until subjects dismissed it with a key press. Then a grey screen was presented for 1 second. Then a prime was presented for 2 seconds. Then a grey screen was presented for 1 second. Then the squares with dots were presented until the subject responded, with a 10 second timeout if they did not respond.

<figure>
<img src="./flow-of-experiment4.pdf" alt="Mountain View" style="width:480px;height:300px;">
<figcaption>Flow of experiment 4.</figcaption>
</figure>

# Describe the data

Here is a description of each column in the raw data.

* **Sub** identifier for subject. There are 40 subjects.
* **Trl** identifier for the trial. There are 64 Trials for each subject.
* **Itm** identifier for item. There are 4 items. An item is a set of Selections for the dots in the squares, e.g., Item 1 is (6,15,24) in any order. This should be included as a (possibly ordered) factor because the task will probably take longer for more dots. Include this in the final model if it does increase rt. Table below shows the dots for each item.
* **Cnd** just a single vector for the combination of vagueness and selection. Useful for some plotting routines that don't accept pairs of grouping variables. 
* **Ord** whether the order was for example (6,15,24) or (24,14,6). This matters because when the smaller quantity is on the left people are quicker than when the smaller quantity is on the right (i.e, SNARC effect). It is a control variable rather than a direct predictor. It could be left out of the main model assuming it is non-significant.
* **Qty** whether the instruction identifies bigger or smaller quantities. This is a factor with 2 levels sum coded. An example of bigger is _"Choose a square with more dots than the target"_. An example of smaller is _"Choose a square with fewer dots than the target"_.
* **Prm** how many dots there were in the prime (prime is _target_ in the instructions to the participants). Primes were 6 10 16 20 24 26 30 34 36 40 44 50 54.
* **Lft** how many dots there were in the left square
* **Mid** how many dots there were in the middle square
* **Rgt** how many dots there were in the right-hand side square
* **key** which key the subject pressed as a response on a trial.
* **rt** raw response time in milliseconds
* **Ins** a character array for the instruction that was presented to the subject on that trial. Table below shows an example of an instruction in each condition.
* **V** factor for vagueness with 2 levels sum coded
* **S** factor for selection task with 2 levels sum coded

Here is a table of instructions.

<style type="text/css">
.table {width: 50%;}
</style>

vagueness | selection | instruction
----------|-----------|--------------------------------------------------------------------
crisp     | match     | Choose a square with the same number of dots as the target  
crisp     | compare   | Choose a square with fewer dots than the target 
vague     | match     | Choose a square with about the same number of dots as the target 
vague     | compare   | Choose a square with far fewer dots than the target

Item   | dots
-------|-----------
Item 1 | (6,15,24)
Item 2 | (16,25,34)
Item 3 | (26,35,44)
Item 4 | (36,45,54)

```{r, docOpts, echo=FALSE, cache=FALSE}
library(knitr)
knit_hooks$set(crop=hook_pdfcrop)
```

```{r, loadLibraries, echo=FALSE}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(latticeExtra))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

Source additional R files with functions for computing confidence intervals.

```{r, sourceFunctions}
source("CIfunctions.R")
```

Get data from the raw per-subject files if necessary.

```{r, getData}
if (file.exists('data.txt')) {
	dat <- read.delim(file="data.txt", header=T, stringsAsFactors=FALSE)
} else {
	system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "data.txt", sep=""))
	system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "data.txt", sep=""))
}

# declare local variables
Selection_of_valid_subjects <- 40
Selection_of_rows <- 7680
Selection_of_trials_per_subject <- Selection_of_rows / Selection_of_valid_subjects # 192

# sort out borderline responses into expected near and far
# How many squares in the square they chose? 
dat$RESPONSE <- as.character(dat$key)
for (row in 1 : nrow(dat) ) {
  switch(dat[row,'RESPONSE'],
         'L' = {dat[row, 'choice'] <- dat[row, 'Lft']},
         'M' = {dat[row, 'choice'] <- dat[row, 'Mid']},
         'R' = {dat[row, 'choice'] <- dat[row, 'Rgt']}         
  )
}
dat$Item <- dat$Itm
dat$Itm <- NULL
dat$crossed=paste('Con',dat$Cnd,':Quan',dat$Qty,':Item',dat$Item,sep="")

dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36

dat$isResponseExpected <- dat$choice == dat$ResponseExpected
dat$isResponseNear <- dat$choice == dat$ResponseNear
dat$isResponseFar <- dat$choice == dat$ResponseFar

for ( row in 1:nrow(dat) ) {
  dat[row, 'switchResponse'] <- 
    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
}

# SUBJECT
# ensure Subject is a factor 
dat$Subject=factor(paste("s",sprintf("%02d",dat$Sub),sep=""))

# TRIAL
# Trial for subject, 1 to 192
dat$Trial = rep(x = 1:Selection_of_trials_per_subject, times = Selection_of_valid_subjects)

# make a centred Trial for modeling
dat$c_Trl <-dat$Trial - mean(dat$Trial)

# make a scaled Trial for modelling
dat$s_Trl <- as.numeric(scale(dat$Trial))

# ID
# id is a unique identifier for the 7680 row data
dat$id <- factor(paste(paste(dat$Subject), 
                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

# ITEM
# create a centred numeric item variable for modeling
dat$c_Itm <- ifelse(dat$Item==1, -.75, 
                    ifelse(dat$Item==2, -.25,
                           ifelse(dat$Item==3, .25, .75)))

# make Item be a factor and assign labels
dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))

# add ratios for Item
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
# 0.2500000 0.4705882 0.5909091 0.6666667
item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
# -1.3441995 -0.1316642  0.5297187  0.9461450
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
# 0.5125000 0.6876471 0.7691558 0.8166667
item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
# -1.37582241 -0.06614191  0.54334858 0.89861574

dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667

dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450

dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667

dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

# VAGUENESS
# Create a factor coding for Vagueness
dat[ dat$Cnd==1 , 'Vagueness'] <- 'Vague'
dat[ dat$Cnd==2 , 'Vagueness'] <- 'Crisp'
dat[ dat$Cnd==3 , 'Vagueness'] <- 'Vague'
dat[ dat$Cnd==4 , 'Vagueness'] <- 'Crisp'
dat$Vagueness <- as.factor(dat$Vagueness)

# manually center Vagueness
dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)

dat$Selection = ""
dat[dat$Cnd %in% c(1,2), "Selection"] <- "matching"
dat[dat$Cnd %in% c(3,4), "Selection"] <- "comparison"
dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
dat$Selection <- as.factor(dat$Selection)

# remove redundant confusing columns
dat$Sub <- NULL
dat$Trl <- NULL

# ORDER
# give the levels of Order meaningful names
dat$Order <- factor(dat$Ord, levels=c(1,2), labels=c('LtoR','RtoL'))

# make a manually centred Order
dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)

# QUANTITY
# give the levels of Quantity meaningful names
dat$Quantity <- factor(dat$Qty, levels=c(1,2), labels=c('Small','Large'))

# make a manually centred Quantity
dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)

# INSTRUCTION
# add Selection of characters in the instruction # 29 30 34 36 38
dat$nchar_instr = nchar(as.character(dat$Ins))
dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Ins)), scale=TRUE))

# make Instruction be a factor (17 levels)
dat$Ins <- as.factor(dat$Ins)

# RT
# add transformations of RT
dat$RT <- dat$rt; dat$rt <-  NULL
dat$RT_RecSqd    <- 1/dat$RT^2
dat$RT_Rec       <- 1/dat$RT
dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
dat$RT_log       <- log(dat$RT)
dat$RT_sqt       <- sqrt(dat$RT)
dat$RT_raw       <- dat$RT
dat$RT_sqd       <- dat$RT^2

# lose impossible trials
dd <- dat
dd$RT[dd$RT > 9000 ] <- NA
dd <- dd[complete.cases(dd),]
row.names(dd) <- NULL

# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
dd$RTprev <- NA
for (s in levels(dd$Subject)) {
  nrows = nrow(dd[dd$Subject==s,])
  for (i in 1:nrows) {
    if (i==1) {
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
    }
    else
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
  }
}

# add transformations of previous RT
dd$RTprev_RecSqd    <- 1/dd$RTprev^2
dd$RTprev_Rec       <- 1/dd$RTprev
dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
dd$RTprev_log       <- log(dd$RTprev)
dd$RTprev_sqt       <- sqrt(dd$RTprev)
dd$RTprev_raw       <- dd$RTprev
dd$RTprev_sqd       <- dd$RTprev^2

save(dd, file="data.Rda")
```

@@

```{r foo,fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

@@

# create extra rows for dfs with removed (outlier) data points
dat1 <- dat_transforms                                       # nrow 7680
dat2 <- subset(dat_transforms, RT<9000)              # nrow 7677
dat2$dfSubset <- "RT<9000"

# bring the dfs back into main df
dat_transforms <- rbind(dat1,dat2)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT<9000"))

@@

temp <- reshape2::melt(data=dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"))

ggplot(temp, 
			 aes(value, colour=RTtype)) + 
	geom_freqpoly() + 
  facet_wrap(dfSubset~variable, scales="free", nrow=2) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r bar, fig.width=20, fig.height=5}
# boxplot subjects and items
load("data.Rda")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw")) 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT.

```{r, bar2, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups=si$lsign,
              xlab=NULL,
              ylab="RT (in ms)",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from mean RT (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

Show how practice and fatigue effects vary over subjects

```{r barr3, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

Show practice and fatigue effects vary over items.

```{r, var3, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

# Show means.

Main effects

```{r, bar4}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

# plot main effects
ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r, bar5, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.01) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
```

2 way interactions ignoring item

```{r, bar6}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Selection",4)
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Selection:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Selection:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
```

```{r,bar7,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

>>>>

@@

```{r,fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("ggplot2"))
load("dat.Rda")

@@

# create extra rows for dfs with removed (outlier) data points
dat1 <- dat_transforms                                       # nrow 7680
dat2 <- subset(dat_transforms, RT>1)                         # nrow 7678
dat2$dfSubset <- "RT>1"
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"

# bring the dfs back into main df
dat_transforms <- rbind(dat1, dat2, dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT>1", "RT>1&RT<59999", "RT>1&RT<30001"))

@@

temp <- melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
             variable.name="transformation",
             value.name="score")

ggplot(temp, aes(score, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(dfSubset~transformation, scales="free", nrow=4) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```
